Automatically Proving Equivalence by Type-Safe Reflection

### 2017-03-31 paper-CICM-17/content/1-introduction.tex
Restored the changes from Edwins two commits

When reading these definitions, one could wrongly think that the propositional equality is a simple wrapper of the judgemental equality. This is not the case : the propositional equality does not only contain the judgemental equality, because in these type theories, a principle of induction is associated with each inductive type. If a type $T$ is an inductive type with a constant constructor and a recursive constructor, ie, $T = 1 + T$, defined in Idris as :

@@

We revisit an example from a previous work of Edwin Brady~\cite{DBLP:conf/plpv/Brady13} which shows how proof obligations may arise when a type is indexed by natural numbers. In this development, the goal is to implement a formally verified library of binary numbers. In order to ensure the correctness of the methods that will be developed, we define the types \code{Bit} and \code{Binary} as indexed over the value they represent (expressed as a natural number), following a correct-by-construction approach.

>>>>

When reading these definitions, one could think that the propositional equality is a simple wrapper of the judgemental equality. This is not the case : the propositional equality does not only contain the judgemental equality, because in these type theories, a principle of induction is associated with each inductive type. If a type $T$ is an inductive type with a constant constructor and a recursive constructor, ie, $T = 1 + T$, defined in Idris as :

@@

We revisit an example from previous work~\cite{DBLP:conf/plpv/Brady13} which shows how proof obligations may arise when a type is indexed by natural numbers. In this development, the goal is to implement a formally verified library of binary numbers. In order to ensure the correctness of the methods that will be developed, we define the types \code{Bit} and \code{Binary} as indexed over the value they represent (expressed as a natural number), following a correct-by-construction approach.

### 2017-03-31 paper-CICM-17/content/3-hierarchyOfTactics.tex
Restored the changes from Edwins two commits

If the user wants to prove propositional equalities, then he will simply instantiate $(\simeq)$ with the built-in $(=)$ during the definition of the \code{Set} implementation.
Note that $(\simeq)$ is only weakly decidable in the sense that \code{set\_eq} only produces a proof when the two elements are equivalent, but it doesn't produce a proof of dis-equivalence when they are different -- instead, it simply produces the value \code{Nothing}--. That is normal, as we want to generate proof of equivalence, and not to generate counter examples for proving dis-equivalence, which is another problem.

Obviously, there is no tactic associated to \code{Set}, since we have no operations and no properties associated to this structure. Therefore, equivalences in a \code{Set} are just the ``syntactic equivalences", and they can simply be proven with \code{refl}\footnote{\code{refl} should not be confused with \code{Refl}, the only constructor of $=$. But of course, when $(\simeq)$ is instantiated with the equality $=$, the instantiation of \code{refl} will be \code{Refl}. Therefore, \code{refl} of the interface \code{Set} is a generalisation of the most specialised \code{Refl} for \code{=}}.

@@

For an expression $e_x\ :\ \code{ExprMa}\ \Gamma\ x$, we will say that $e_x$ denotes (or encodes) the number $x$ in the context $\Gamma$.
When an expression is a variable \code{(VarMa\ \_\ $\Gamma$\ i)}, the denoted number is simply the corresponding variable in the context, ie, \code{(index\ i\ $\Gamma$)}.
The expression \code{(ConstMa\ \_\ $\Gamma$\ k)} denotes the constant $k$ of type $c$. Finally, if $e_x$ is an expression encoding the concrete value $x$, and $e_y$ is an expression encoding the concrete value $y$, then the expression \code{PlusMa\ $e_x$\ $e_y$} denotes the concrete value $(x + y)$. Because the reflected terms embed their corresponding inputs, they are guarantee to be faithful representations of them.

@@

Finally, the main function which tries to prove the equivalence $x \simeq y$ simply has to normalise the two reflected terms encoding the left and the right hand side, and to use the function \code{buildProof} in order to compose the two proofs that have been obtained :

>>>>

If the user wants to prove propositional equalities, they will instantiate $(\simeq)$ with the built-in $(=)$ during the definition of the \code{Set} implementation.
Note that $(\simeq)$ is only weakly decidable in the sense that \code{set\_eq} only produces a proof when the two elements are equivalent, but it doesn't produce a proof of dis-equivalence when they are different -- instead, it simply produces the value \code{Nothing}--. That is normal, as we want to generate proof of equivalence, and not to generate counter examples for proving dis-equivalence, which is another problem.

There is no tactic associated to \code{Set}, since we have no operations and no properties associated to this structure. Therefore, equivalences in a \code{Set} are just the ``syntactic equivalences", and they can simply be proven with \code{refl}\footnote{\code{refl} should not be confused with \code{Refl}, the only constructor of $=$. But of course, when $(\simeq)$ is instantiated with the equality $=$, the instantiation of \code{refl} will be \code{Refl}. Therefore, \code{refl} of the interface \code{Set} is a generalisation of the most specialised \code{Refl} for \code{=}}.

@@

For an expression $e_x\ :\ \code{ExprMa}\ \Gamma\ x$, we will say that $e_x$ denotes (or encodes) the number $x$ in the context $\Gamma$.
When an expression is a variable \code{(VarMa\ \_\ $\Gamma$\ i)}, the denoted number is the corresponding variable in the context, ie, \code{(index\ i\ $\Gamma$)}.
The expression \code{(ConstMa\ \_\ $\Gamma$\ k)} denotes the constant $k$ of type $c$. Finally, if $e_x$ is an expression encoding the concrete value $x$, and $e_y$ is an expression encoding the concrete value $y$, then the expression \code{PlusMa\ $e_x$\ $e_y$} denotes the concrete value $(x + y)$. Because the reflected terms embed their corresponding inputs, they are guarantee to be faithful representations of them.

@@

Finally, the main function which tries to prove the equivalence $x \simeq y$ has to normalise the two reflected terms encoding the left and the right hand side, and to use the function \code{buildProof} in order to compose the two proofs that have been obtained :
